---
title: "Exchange Rates Networks"
author: "Thorsten Simon, Payton J. Jones, Patrick Mair, and Achim Zeileis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Exchange Rates Networks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{networktree}
  %\VignetteKeywords{regression trees, networks, exchange rates}
  %\VignettePackage{networktree}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library("networktree")
# library("zoo")
set.seed(111)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##"
)
```

## Data

We use the `FXRatesCHF` data from package **fxregime** [@fxregime],
which is a  zoo series containing 25 daily time series over four decades.
The columns correspond to the prices for various currencies w.r.t. CHF
(Swiss Franc)
as unit currency.

Here, we aim at investigating the network of weekly returns for
seven currencies:

| Abbreviation | Full name  |
|:---|:--------------------------------------------------------------------|
| USD  | US Dollar |
| JPY  | Japanese Yen |
| DUR  | A mixture of the German Mark (DEM, before 1999) and the Euro (EUR, since 1999)   |
| CNY  | Chinese Yuan |
| INR  | Indian Rupee |
| GBP  | Great British Pound |
| ZAR  | South African Rand |


```{r data}
data("FXRatesCHF", package = "fxregime")
currencies <- c("USD", "JPY", "DUR", "CNY", "INR", "GBP", "ZAR")
returns <- fxregime::fxreturns(currencies, data = FXRatesCHF)
returns <- na.omit(returns)
```

```{r look-at-data, fig.height = 7, fig.width = 12, out.width = "100%"}
plot(returns)
```

A quick look reveals that data are on hand from `r head(zoo::index(returns), 1)`
to `r tail(zoo::index(returns), 1)`. Over the whole period the mean of each time
series is around zero. Further the spread for each time series is relatively
constant over time. However, the correlation structure between these seven
variables might change within this period of time.

Before modelling the _networktree_ we compute a numeric variable for the
decimal time and put it in a column `time`, and for illustration we also
add a nosie variable `foo`. Obviously, one can expect that `time` will
be selected for splitting the data, but `foo` will not be selected.

```{r data2}
returns$year   <- as.POSIXlt(zoo::index(returns))$year
returns$isLeap <- (returns$year %% 4) == 0
returns$yday   <- as.POSIXlt(zoo::index(returns))$yday
returns$time   <- 1900 + returns$year + returns$yday / ifelse(returns$isLeap, 366, 365)
returns$foo    <- rnorm(nrow(returns))
```

## Networktree Model

Then we specify the formula object for our _networktree_. On the left hand side we put
the variables among which the correlation structure has to be investigated,
i.e. the currencies. On the right hand side the split variables are placed. Here,
`time` and `foo`.

```{r formula}
f  <- paste(paste(currencies, collapse = " + "), "~ time + foo")
f  <- as.formula(f)
f
```

We put this formula into the `networktree()` function, where we also
specify that only the correlations should be considered while splitting `cor = TRUE`
(not the means and standard deviations). Further the tree should only grow to a maximum
depth of three `maxdepth = 3`. The default method for the recursive partitioning
is the _model-based recursive partitioning_ [@mob] implemented in the
function `mob()` of the **partykit** package [@partykit].

```{r model}
tr <- networktree(f, data = returns, cor = TRUE, maxdepth = 3)
```

```{r output}
print(tr, digits = 2)
```

The `print()` method gives the partitioning of the data and the fitted coefficients
which are the correlations between the exchange return rates of the different
currencies. As expected only `time` was selected for splitting the data.
In the first node the data are splitted in
`r format(zoo::as.yearmon(tr[[1]]$node$split$breaks), "%B %Y")`.
Then in the second and fifth node the remaining data is splitted in
`r format(zoo::as.yearmon(tr[[2]]$node$split$breaks), "%B %Y")` and
`r format(zoo::as.yearmon(tr[[5]]$node$split$breaks), "%B %Y")`,
respectively.


```{r output2, fig.align = "left", fig.height = 16, fig.width = 24, dpi = 72, out.width = "100%"}
plot(tr, type = "pcor", posCol = "#008585", negCol = "#C7522B", maximum = 1)
```

The `plot()` method uses `qgraph()` of the **qgraph** package [@qgraph] for
visualizing the networks in the terminal nodes.  Thus we can choose between
looking at the correlations `type = "cor"` or at the partial correlations
`type = "pcor"`.

The partial correlation reveals a very strong link between the Chinese Yuan and
the US Dollar after April 1994, but a weaker link before that point of time.

The split which was found in the second part of the data was July 2005
which coincide with an announcement of the People's Bank of China (PBC)
to shift away from a fixed exchange rate to the US Dollar to a basket
of currencies with greater flexibility [@zeileis2010].
Earlier studies also found that CNY was still strongly coupled to the
USD after July 2005 [@zeileis2010].
However, the split detected in mid 2005 by the _networktree_ does not show
a big change in the coupling of CNY to USD, but reveals that coupling of
Indian Rupee, Chinese Yuan, and South African Rand to the Euro increased.

The split in April 1994 raises the question why the coupling of CNY to USD
was that weak before that point of time.

<!--
```{r investigate-split, fig.align = "left", fig.height = 9, fig.width = 9, dpi = 144, out.width = "100%"}
# partykit::nodeapply(tr, ids = c(3, 4, 6, 7),
#   FUN = function(n){ plot(CNY ~ USD, data = n$data) })
# 
# partykit::nodeapply(tr, ids = c(3, 4, 6, 7),
#   FUN = function(n){ summary(lm( CNY ~ USD, data = n$data ))$r.squared })

par(mfrow = c(2, 2))
for (i in partykit::nodeids(tr, terminal = TRUE)) {
  sfm <- summary(fm <- lm(CNY ~ USD, data = tr[[i]]$data))
  plot(CNY ~ USD, data = tr[[i]]$data, main = sprintf("node %d", i))
  abline(fm, col = "darkred")
  legend("topleft", bty = "n",
    legend = sprintf("b0 = %.3f (%.3f)\nb1 = %.3f (%.3f)\nR2 = %.3f",
                     sfm$coefficients[1, 1], sfm$coefficients[1, 2],
                     sfm$coefficients[2, 1], sfm$coefficients[2, 2],
                     sfm$r.squared))
  
}
```
-->

## License

This content is published under [Creative Commons Attribution-NonCommercial-ShareAlike 4.0
International (CC BY-NC-SA 4.0)](https://creativecommons.org/licenses/by-nc-sa/4.0/).

## References


